### 2020-07-14 steadystate.tex
final comment edits

\section{Steady-State Thermal-Hydraulics Results}

Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
fuel salt in the core at steady state from Moltres and the Polimi and TUDelft
models. Figure \ref{fig:flow} provides an alternate view of the flow profile
through flow streamlines superimposed on the velocity magnitude distribution.
The results from Moltres show good qualitative agreement with the Polimi and
TUDelft models \cite{fiorina_modelling_2014}; the plots show similar
flow and hotspot features in all three models. Furthermore, the highest salt
velocities in all three models occur at the inlet, outlet, and at core
half-height approximately 0.40 m away from the central axis. A large
recirculation region forms near the blanket tank walls from turbulent flow.
Inertial forces dominate over viscous forces to form this large eddy. The main
difference between Moltres and Fiorina et al.'s models is the flow profile
near the central axis at the top and bottom of the core.
The Polimi and TUDelft models predict relatively stagnant flow in these
regions without recirculation. Moltres, on the other hand, predicts explicit
recirculating flow in these regions. This is due to our constant turbulent
viscosity approximation in Moltres. The k-$\epsilon$ turbulence models in the
Polimi and TUDelft models predict that the turbulent viscosity in these
regions is as high as 100 Pa$\cdot$s, much higher than our 40 Pa$\cdot$s
approximation.

@@

Figure \ref{fig:pre} compares the total delayed neutron source
distribution from Moltres with the results from the Polimi and TUDelft models
\cite{fiorina_modelling_2014}. In contrast to Figure \ref{fig:dnp} which shows
the precursor distribution, Figure \ref{fig:pre} shows the rate of delayed
neutron emission, which was calculated by multiplying each \gls{DNP} group
$C_i$ with its associated decay constant $\lambda_i$.
The Polimi and TUDelft models
feature greater \gls{DNP} retention in the stagnant regions within the core.
This effect is less pronounced in the Moltres model. Less build-up
of \glspl{DNP} at the top of the core occurs in Moltres because most of the
\glspl{DNP} produced near the center of the core cannot enter the axial
recirculation zone that only appears in Moltres. The figure also shows less
build-up in the large recirculation zone near the blanket tank.

The in-core delayed neutron fraction $\beta_c$ is an important safety
parameter for \glspl{MSR}. This value represents the actual delayed
neutron fraction in \glspl{MSR} after accounting for the loss of delayed
neutrons from \glspl{DNP} decaying outside the active core region.
Reactors with smaller $\beta$ values exhibit greater prompt jumps in the
neutron flux in response to reactivity insertions because they have a greater
proportion of prompt neutrons under normal operating conditions. This is
undesirable from a reactor safety perspective because it exposes the reactor
to relatively more extreme conditions before active safety mechanisms
activate and scram the reactor. In \glspl{MSR}, this danger is partly
mitigated by the strong, negative fuel temperature reactivity coefficient. 
Chapter 6 contains more in-depth discussions for various transient scenarios.

Table \ref{table:dnf} compares the fraction of out-of-core emissions and
the $\beta_c$ values from Moltres with the Polimi and TUDelft models. We
calculated the fraction of out-of-core emissions by finding the total amount
of each \gls{DNP} group in the core and the outer loop, multiplying each total
by their associated decay time constants $\lambda_i$, and calculating the
proportion of emissions in the outer loop relative to the grand total. We
calculated $\beta_c$ by first obtaining the prompt neutron emission rate from
Moltres and subsequently using in-core delayed neutron emission rate from the
previous calculation to find the fraction of delayed neutron emission rate
relative to total emission rate.

@@

The fraction of out-of-core emissions from our Moltres model
differs significantly by approximately 10\%, and $\beta_c$ differs by
60-70 pcm. The former is attributed to the lesser \gls{DNP} retention in the
stagnant flow regions in the core; the \glspl{DNP} are more evenly distributed
along the entire primary loop, leading to more delayed neutron emissions in
the outer loop region. This is likely caused by differences in the
flow pattern in the recirculation zone because convective transport
dominates diffusive effects in the \gls{MSFR}. The exact flow pattern in the
recirculation zones in the Polimi and TUDelft models is likely to differ from
that in Moltres (Figure \ref{fig:flow}). Figure \ref{fig:flow-temp} also shows
some minor differences in the magnitude of the flow in the recirculation zones
between Moltres, and the Polimi and TUDelft models. Although the sizes of the
arrows representing flow velocity are on different scales, a
quick comparison between the largest arrows and the arrows in the
recirculation zone indicates that the recirculation zone in Moltres is
relatively more stagnant. This could explain the concentration of \glspl{DNP}
along an ``arc'' closer to the center of the core in Moltres as opposed to the
more even distribution of \glspl{DNP} throughout the whole recirculation zones
in the Polimi and TUDelft models. The higher peak \glspl{DNP} distribution in
Moltres also supports this assertion.

In spite of the greater delayed neutron losses, the $\beta_c$ value is higher
in Moltres than the Polimi and TUDelft models. To account for this
peculiarity, Fiorina et al. \cite{fiorina_modelling_2014} applied adjoint flux
weighting for their $\beta_c$ calculation whereas this work
reports the value as the physical fraction without adjoint weighting. The
weighting results in a significant difference in $\beta_c$ because a greater
fraction of the \glspl{DNP} decay in the recirculation zones, where the
neutron importance is noticeably diminished.

@@

The movement of decay heat precursors effectively redistributes a fraction of
the volumetric heat source concentrated at the center of the core to the
entire primary loop.
Thus, the temperature distribution in the core should
show less extreme temperatures compared with the simulation without decay
heat.

Figure \ref{fig:decayheat} shows the decay heat source distribution in the
core for groups 1 to 3. The decay constants are $\lambda_j = 0.197$ s$^{-1},
0.0168$ s$^{-1}, 0.000358$ s$^{-1}$ for groups 1, 2, and 3, respectively.
Similar to the delayed neutron precursor distributions, the shorter-lived
decay heat precursors show greater build-up in the recirculation zones
compared with the longer-lived decay heat precursors that are more evenly
distributed throughout the core.

@@

Figure \ref{fig:decayheattemp} shows the
difference in core temperatures at steady state with decay heat modeling
relative to the core temperatures without decay heat in Figure
\ref{fig:flow-temp}. This figure uses a different colormap scheme to clearly
distinguish between temperature increases and decreases in the core. The
hotspots in the recirculation zones are approximately
2 to 4 K cooler, while the cooled salt flowing in from the inlet is
approximately 2 K hotter. These regions correspond to the hottest and coldest
areas in the core and the results agree with the expectation that the extreme
temperatures would be affected the most by the introduction of decay heat
precursors. The difference between the outlet and inlet temperatures falls
from 101.1 K to 98.9 K as some of the decay heat precursors deposit heat in
the outer loop region. The figure also shows some unphysical oscillations in
the temperature near the outlet because the mesh near the boundary is too
coarse. Users can avoid these oscillations near the outlet but using a mesh
that is progressively finer the closer the element is to the boundary. The
present author did not resolve this issue due to time constraints.

>>>>

\section{Steady-State Thermal-Hydraulics Results} \label{sec:ss-th}

Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
fuel salt in the core at steady state from Moltres and the Polimi and TUDelft
models. Figure \ref{fig:flow} provides an alternate view of the flow profile
through flow streamlines superimposed on the velocity magnitude distribution.
The results from Moltres show good qualitative agreement with the Polimi and
TUDelft models \cite{fiorina_modelling_2014}; the plots show similar
flow and hotspot features in all three models. Furthermore, the highest salt
velocities in all three models occur at the inlet, outlet, and at core
half-height approximately 0.40 m away from the central axis. A large
recirculation region forms near the blanket tank walls from turbulent flow.
Inertial forces dominate over viscous forces to form this large eddy. The main
difference between Moltres, and the Polimi and TUDelft models is the flow
profile near the central axis at the top and bottom of the core.
The Polimi and TUDelft models predict relatively stagnant flow in these
regions without recirculation. Moltres, on the other hand, predicts explicit
recirculating flow in these regions. This is due to our constant turbulent
viscosity approximation in Moltres. The k-$\epsilon$ turbulence models in the
Polimi and TUDelft models predict that the turbulent viscosity in these
regions is as high as 100 Pa$\cdot$s, much higher than our 40 Pa$\cdot$s
approximation.

@@

Figure \ref{fig:pre} compares the total delayed neutron source
distribution from Moltres with the results from the Polimi and TUDelft models
\cite{fiorina_modelling_2014}. In contrast to Figure \ref{fig:dnp} which shows
the precursor distribution, Figure \ref{fig:pre} shows the rate of delayed
neutron emission, which was calculated by multiplying each \gls{DNP} group
$C_i$ with its associated decay constant $\lambda_i$.
The Polimi and TUDelft models
feature greater \gls{DNP} retention in the stagnant regions in the core. This
effect is less pronounced in the Moltres model, most notably at the top of the
core near the central axis and in the recirculation zone adjacent to the
blanket tank. As mentioned in Section \ref{sec:ss-th}, only the Moltres model
features a small recirculation zone at the top of the core. This local flow
pattern is responsible for diverting the shorter-lived \glspl{DNP} away from
the top of the core and reducing the build-up in that region.

The in-core delayed neutron fraction, which we denote as $\beta_c$, is an
important safety parameter for \glspl{MSR}. This value represents the actual
delayed neutron fraction in \glspl{MSR} after accounting for the loss of
delayed neutrons from \glspl{DNP} decaying outside the active core region.
Reactors with smaller $\beta$ values exhibit greater prompt jumps in the
neutron flux in response to reactivity insertions because they have a greater
proportion of prompt neutrons under normal operating conditions. This is
undesirable from a reactor safety perspective because it reduces the time
available for active safety mechanisms to
activate and scram the reactor. In \glspl{MSR}, this danger is partly
mitigated by the strong, negative fuel temperature reactivity coefficient. 
Chapter 6 contains more in-depth discussions for various transient scenarios.

Table \ref{table:dnf} compares the fraction of out-of-core emissions and
the $\beta_c$ values from Moltres with the Polimi and TUDelft models. The
present author calculated the fraction of out-of-core emissions by finding the
average concentration of each \gls{DNP} group in the core and the outer loop,
multiplying each average by their associated decay time constants $\lambda_i$
for the average emission rate, and calculating the
proportion of emissions in the outer loop relative to the total sum. The
present author calculated $\beta_c$ by first obtaining the prompt neutron
emission rate from Moltres and subsequently using the in-core delayed neutron
emission rate from the previous calculation to find the fraction of delayed
neutron emission rate relative to total emission rate.

@@

The fraction of out-of-core emissions from our Moltres model
differs significantly by approximately 10\%, and $\beta_c$ differs by
60-70 pcm. The former is attributed to the lesser \gls{DNP} retention in the
stagnant flow regions in the core; the \glspl{DNP} are more evenly distributed
along the entire primary loop and a greater fraction of delayed neutron
emissions occur in the outer loop region. The flow patterns are responsible
for the distribution of shorter-lived \gls{DNP} since convection is the
dominant mode of species transport in the \gls{MSFR}. The exact flow pattern
in the recirculation zones in the Polimi and TUDelft models differs from
that in Moltres (Figure \ref{fig:flow}). Figure \ref{fig:flow-temp} also shows
some minor differences in the magnitude of the flow in the recirculation zones
between Moltres, and the Polimi and TUDelft models. Although the sizes of the
arrows representing flow velocity are on different scales, a
quick comparison between the largest arrows and the arrows in the
recirculation zone indicates that the recirculation zone in Moltres is
more stagnant. This could explain the concentration of \glspl{DNP}
along an ``arc'' closer to the center of the core in Moltres as opposed to the
more even distribution of \glspl{DNP} throughout the whole recirculation zones
in the Polimi and TUDelft models. The higher peak \gls{DNP} distribution in
Moltres also supports this assertion.

In spite of the greater delayed neutron losses, the $\beta_c$ value is higher
in Moltres than the Polimi and TUDelft models. Fiorina et al.
\cite{fiorina_modelling_2014} applied adjoint flux
weighting for their $\beta_c$ calculation whereas this work
reports the value as the physical fraction without adjoint weighting. The
weighting results in a significant difference in $\beta_c$ because a greater
fraction of the \glspl{DNP} decay in the recirculation zones, where the
neutron importance is noticeably diminished.

@@

The movement of decay heat precursors effectively redistributes a fraction of
the fission heat source concentrated at the center of the core to other parts
of the primary loop. Thus, the temperature distribution in the core should
show less extreme temperatures compared with the simulation without decay
heat.

Figure \ref{fig:decayheat} shows the decay heat source distribution in the
core for groups 1 to 3. The decay constants are $\lambda_j = 0.197$ s$^{-1},
0.0168$ s$^{-1}, 0.000358$ s$^{-1}$ for decay heat groups $j=$ 1, 2, and 3,
respectively.
Similar to the delayed neutron precursor distributions, the shorter-lived
decay heat precursors show greater build-up in the recirculation zones
compared with the longer-lived decay heat precursors that are more evenly
distributed throughout the core.

@@

Figure \ref{fig:decayheattemp} shows the
difference in core temperatures at steady state with decay heat modeling
relative to the core temperatures without decay heat in Figure
\ref{fig:flow-temp}. This figure uses a different color map to clearly
distinguish between positive and negative temperature differences in the core.
The hotspots in the recirculation zones are approximately
2 to 4 K cooler, while the cooled salt flowing in from the inlet is
approximately 2 K hotter. These regions correspond to the hottest and coldest
areas in the core and the results agree with the expectation that the extreme
temperatures would be affected the most by the introduction of decay heat
precursors. The difference between the outlet and inlet temperatures falls
from 101.1 K to 98.9 K as some of the decay heat precursors deposit heat in
the outer loop region. The figure also shows some unphysical oscillations in
the temperature near the outlet because the mesh near the boundary is too
coarse. Users can avoid these oscillations near the outlet by using a mesh
that is progressively finer towards the boundary. The
present author could not resolve this issue due to time constraints.
